---
title: 'Stat 579 - Homework #4'
author: "Your Name"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsets and Visualizations: Movie Mojo

For this homework we are using the data set `mojo` from the classdata package. `mojo` constains data on box office revenue for movies based on the website https://www.boxofficemojo.com .


Run the following two commands to install the newest version of the package from github:

```{r, eval = FALSE}
library(devtools) # error? then run install.packages("devtools")
devtools::install_github("heike/classdata")
```

Check what is in the data set:
```{r}
library(classdata)
head(mojo)
library(ggplot2)
```

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.

2. What is the difference between the variables `Week` and `WeekNo`? Describe in your words.

```{r}
str(mojo)
summary(mojo$Week)
summary(mojo$WeekNo)
boxplot(mojo$Week,main="Boxplot of Week", ylab="Week")
boxplot(mojo$WeekNo,main="Boxplot of WeekNo", ylab="WeekNo")

```

Both Week and WeekNo are of type Num. They differ in values and the range. Week is in range -48 to 873 whereas, WeekNo is in range 1 and 53. It seems like WeekNo is the week in a given year. Further I found, using `?mojo`

Week week that the movie has been in theaters
WeekNo integer, week number of the year (1 through 52 or 53)

3. Use `ggplot2` to plot total gross (`Total Gross`) against week number (`WeekNo`). Facet by Year. Interpret the result. 
Which movie had the highest total gross over the time frame? How many weeks was that movie on rank 1? How long was it in box offices overall?

```{r}
cap<-paste(strwrap("Plot showing total gross against WeekNo. For each week of the year multiple movies have different total gross values. The highest total gross value is reached in Week 24 of year 2016.",width = 95),collapse = "\n")
ggplot(data = mojo,aes(x=WeekNo,y=`Total Gross`))+ geom_point()+ facet_wrap( ~ Year) +labs(caption=cap)

#highest total gross
hg<-mojo$Title[which.max(mojo$`Total Gross`)]
#all data about highest total gross
mojo[which.max(mojo$`Total Gross`),]
#weeks in theater
mojo$Week[which.max(mojo$`Total Gross`)]

#find 1 weeks
length(which(mojo$Title==hg & mojo$TW==1))
```

Highest total gross was `r mojo$Title[which.max(mojo$'Total Gross')]` it was in box office for `r mojo$Week[which.max(mojo$'Total Gross')]` weeks. This movie was number 1 for `r length(which(mojo$Title==hg & mojo$TW==1))` weeks.


4. Pick two movies that were in box office some time between 2013 and 2018 and find the corresponding data in the `mojo` data. How does total gross of the two movies compare? Draw a plot and comment on the result. 

```{r}
#total movies in mojo
length(mojo$Title[which(mojo$Year>=2013 & mojo$Year<=2018)])
# select two movies from mojo
m1<-"Super Troopers 2"
m2<-"A Quiet Place"

#data for movie1
mojo[which(mojo$Title==m1),]
#data for movie2
mojo[which(mojo$Title==m2),]

twoMovies<- rbind(mojo[which(mojo$Title==m1),],mojo[which(mojo$Title==m2),])

cap<-paste(strwrap("Plot showing total gross against Week for the two selected movies. A Quiet Place started earning at 50 million whereas Super Troopers 2 started at 15 million. The trajectory over the weeks show that A Quiet Place earned much more than Super Troopers 2.",width = 95),collapse = "\n")
ggplot(data = twoMovies,aes(x=Week,y=`Total Gross`))+ geom_point()+ facet_wrap( ~ Title) +labs(caption=cap)

```

5. Hit or Flop? The variable `Budget (in Millions)` contains estimated budget numbers for some movies. For how many movies is this information available (careful! trick question - look at what the function `unique` does)? Studios would like to see their budget returned by the opening weekend. What is the percentage of movies for which that happened? How many movies did not have their budget matched in total gross by the third weekend?
For each of these two questions describe your 'plan of attack', i.e. lay out how you go about finding an answer to the question.
Identify one movie, that did not match its budget by week 3. Plot the incurred loss over time. 


```{r}
#find movies with budget
moviesBudget<-unique(mojo$Title[!is.na(mojo$`Budget (in Million)`)])

moviesBudgetData<- mojo[mojo$Title %in% moviesBudget,]

#find movies with less earning that budget in opening week
flopMovies_wk1<-moviesBudgetData$Title[which(moviesBudgetData$Week==1 & moviesBudgetData$`Budget (in Million)`>moviesBudgetData$`Total Gross`/1000000)]
pct_wk1<-length(unique(flopMovies_wk1))/length(moviesBudget)*100

#
flopMovies_wk3<-moviesBudgetData$Title[which(moviesBudgetData$Week==3 & moviesBudgetData$`Budget (in Million)`>moviesBudgetData$`Total Gross`/1000000)]
pct_wk3<-length(unique(flopMovies_wk3))/length(moviesBudget)*100

#movie turbo earning flop
cap<-paste(strwrap("Plot showing total gross against Week for the movie Turbo. The red line shows the buged of the movie. We can see over the weeks the movie failed to earn more than the budget.",width = 95),collapse = "\n")
turboBugdetinMil=unique(moviesBudgetData[moviesBudgetData$Title=="Turbo",]$`Budget (in Million)`)*1000000
ggplot(data = moviesBudgetData[moviesBudgetData$Title=="Turbo",],aes(x=Week,y=`Total Gross`))+ geom_point() + geom_hline(yintercept=turboBugdetinMil, linetype="dashed", color = "red" , size=1)

#a hit movie example
cap<-paste(strwrap("Plot showing total gross against Week for the movie Spy. The red line shows the buged of the movie. We can see after second week the movie earned more than the budget.",width = 95),collapse = "\n")
spyBugdetinMil=unique(moviesBudgetData[moviesBudgetData$Title=="Spy",]$`Budget (in Million)`)*1000000
ggplot(data = moviesBudgetData[moviesBudgetData$Title=="Spy",],aes(x=Week,y=`Total Gross`))+ geom_point() + geom_hline(yintercept=spyBugdetinMil, linetype="dashed", color = "red" , size=1)


```

To answer the question I first made a subset of data which had all the movies with a budget. There were `r length(moviesBudget)` movies with bugdet. Then I found the movies whose total gross was less than the budget in the first week. I found `r length(unique(flopMovies_wk1))` movies, `r pct_wk1`%, did not earn more than budget in the first week. `r length(moviesBudget)-length(unique(flopMovies_wk1))` were able to earn more than budget in the first week.

Then I found the movies whose total gross was less than the budget in the Third week. I found `r length(unique(flopMovies_wk3))` movies, `r pct_wk3`%, did not earn more than budget in the third week.

The trajectory of earning over weeks from release is shown for a movie "Turbo" which failed to earn the budget. Another movie spy earned more than the budget after 2 weeks.


