---
title: "Stat 579 - Life Expectancy"
author: "Urminder Singh"
date: "9/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=4, fig.height=4)
library(tidyverse)
```

# Life Expectancy in Mexico, Canada and the US 

## Step-by-step plan of work

1. Read the data
2. Focus on values of Canada
3. Visualize the values for Canada
4. See whether we can find the problem
5. ... then we'll see ...

## Reading the data 

```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gapminder-5060.csv" # nasty looking, but doesn't make any assumptions about data on your machine
life5060 <- read.csv(url)
head(life5060)
```

## Focus on the values for Canada

```{r}
canada <- life5060 %>% filter(country == "Canada")
head(canada)
```

## Visualize 

Draw a line for the life expectancy in Canada:
```{r}
canada %>% 
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

Clearly there is something wrong with the value in 1957.

## Find the problematic value and fix it

```{r}
canada %>% filter(year == 1957)

canada_fixed <- canada %>% mutate(
  lifeExp = replace(lifeExp, year==1957, 69.96)  
)
```

## visualize the fixed data

```{r}
canada_fixed %>% ggplot(aes(x = year, y = lifeExp)) +
  geom_line()
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US

```{r}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, group=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))
```

## ... on to the stretch goal ... 

Add lines for Mexico and the US, color lines by country and add a legend

```{r, fig.width = 6}
canada_fixed %>% 
  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = life5060 %>% 
              filter(country %in% c("United States", "Mexico")))

```

## Read 70 80 data

```{r}
url <- "https://raw.githubusercontent.com/Stat579-at-ISU/stat579-at-isu.github.io/master/homework/data/gap7080.csv"
life7080 <- read.csv(url)
head(life7080)

#merge two datasets
lifeAll<-rbind(life5060,life7080)

head(lifeAll)
tail(lifeAll)
```

## Plot  life expectancy for the United States, Canada and Mexico in a line chart

```{r, fig.width = 6}
#fix data for canada
canada <- lifeAll %>% filter(country == "Canada")

canada_fixed <- canada %>% mutate(lifeExp = replace(lifeExp, year==1957, 69.96) )

canada_fixed %>%  ggplot(aes(x = year, y = lifeExp, colour=country)) + geom_line() +
  geom_line(data = lifeAll %>% 
              filter(country %in% c("United States", "Mexico")))

```

The above plot shows the life expectancy for the United States, Canada and Mexico in a line chart from year `r min(lifeAll$year)` till `r max(lifeAll$year)`. In the plot we can see that the life expectancy for all three countries has increased over the given period. For any given year, Canada has had the highest life expectancy whereas Mexico has had the lowest.



