---
title: 'Stat 579 - Homework #8'
author: "Urminder Singh"
date: "10/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Star Wars

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. We are back to a survey collected by FiveThirtyEight. This time we are interested in a survey on Star Wars - the accompanying article is published [here](https://fivethirtyeight.com/features/americas-favorite-star-wars-movies-and-least-favorite-characters/)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readr) 
starwars <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/star-wars-survey/StarWars.csv")

# the following lines are necessary to fix the multibyte problem and make proper names
# part of the names:
line1 <- names(starwars)
line2 <- unlist(starwars[1,])
varnames <- paste(line1, line2)
# clean up some of the multibyte characters:
names(starwars) <- enc2native(stringi::stri_trans_general(varnames, "latin-ascii"))

starwars <- starwars[-1,]
head(starwars)
```

3. How many people responded to the survey? How many people have seen at least one of the movies? Use the variable `Have you seen any of the 6 films in the Star Wars franchise? Response` to answer this question. Only consider responses of participants who have seen at least one of the Star Wars films for the remainder of the homework.

```{r}
#how many responded
dim(starwars)[1]
starwars_yes<-starwars[which(starwars$`Have you seen any of the 6 films in the Star Wars franchise? Response` == 'Yes'),]
#people watched sw
dim(starwars_yes)[1]
```

4. Variables `Gender Response` and `Age Response` are two of the demographic variables collected. Use `dplyr` to provide a frequency break down for each variable. Does the result surprise you? Comment. Reorder the levels in the variable `Age Response` from youngest to oldest.

```{r}
starwars_yes %>% count(`Gender Response`)
starwars_yes %>% count(`Age Response`)
#make factor and arrange levels
starwars_yes$`Age Response`<-factor(starwars_yes$`Age Response`, levels=c("18-29", "30-44", "45-60", "> 60", ""))
levels(starwars_yes$`Age Response`)
```


5. Variables 10 through 15 answer the question: "Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film." for each of the films. 



Bring the data set into a long form. Introduce a variable for the star wars episode and the corresponding ranking. 
Find the average rank for each of the films. Are average ranks different between mens' and womens' rankings?
On how many responses are the averages based? Show these numbers together with the averages.

```{r}
starwars_yes[, 10:15]<-sapply(starwars_yes[, 10:15], as.numeric)

xY<-starwars_yes %>% gather(Ep,Rating,colnames(starwars_yes)[10:15])

xY<-xY[which(!is.na(xY$Rating)),] %>% group_by(Ep) %>% mutate(m=mean(Rating))

#MEAN BY MALE AND FEMALE  
meanRatings<-xY[which(!is.na(xY$Rating)),] %>% group_by(Ep,`Gender Response`) %>% summarise_at(vars("Rating"), funs(mean, length))

ggplot(data=meanRatings,aes(x=`Gender Response`)) + geom_boxplot(aes(y=mean))
```


6. R2 D2 or C-3P0? Which of these two characters is the more popular one? Use responses to variables 25 and 26 to answer this question. Note: first you need to define what you mean by  "popularity" based on the available data.  

```{r}
popularity<-starwars_yes[,25:26] %>% gather(Character,pop,colnames(starwars_yes)[25:26])
#remove NA Unfamiliar (N/A)
popularity<-popularity %>% filter(!is.na(pop))%>% filter(!(pop == "Unfamiliar (N/A)"))
#map popularity to scale -2 -1 0 1 2
# very unfavorably Somewhat unfavorably neutral Somewhat favorably Very favorably
unique(popularity$pop)

popularity <- popularity %>% mutate(popInt = case_when(pop=='Very unfavorably' ~-2, pop=='very unfavorably' ~-2, pop=='Somewhat unfavorably' ~-1, pop=='Neither favorably nor unfavorably (neutral)' ~0, pop=='Somewhat favorably' ~1, pop=='Very favorably' ~2))

popularity %>% group_by(Character) %>% summarise_at(vars(popInt),mean) %>% arrange(desc(popInt))

```




7. Popularity contest: which of the surveyed characters is the most popular? use the popularity measure you defined in the previous question to evaluate responses for characters 16 through 29. Use an appropriate long form of the data to get to your answer. Visualize the result. 

```{r}
popularity<-starwars_yes[,16:29] %>% gather(Character,pop,colnames(starwars_yes)[16:29])
#remove NA Unfamiliar (N/A)
popularity<-popularity %>% filter(!is.na(pop))%>% filter(!(pop == "Unfamiliar (N/A)"))
#map popularity to scale -2 -1 0 1 2
# very unfavorably Somewhat unfavorably neutral Somewhat favorably Very favorably
unique(popularity$pop)

popularity <- popularity %>% mutate(popInt = case_when(pop=='Very unfavorably' ~-2, pop=='very unfavorably' ~-2, pop=='Somewhat unfavorably' ~-1, pop=='Neither favorably nor unfavorably (neutral)' ~0, pop=='Somewhat favorably' ~1, pop=='Very favorably' ~2))

meanpopularityAllchars <- popularity %>% group_by(Character) %>% summarise_at(vars(popInt),mean) %>% arrange(desc(popInt))
#rename
meanpopularityAllchars[1,1]<- "Han Solo"


ggplot(data=meanpopularityAllchars,aes(x=`Character`,y=popInt,fill=(Character))) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


