---
title: 'Stat 579 - Homework #2'
author: "Urminder Singh"
date: "9/12/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Swiss Banknotes

1. Download the RMarkdown file with these homework instructions to use as a template for your work.
Make sure to replace "Your Name" in the YAML with your name.
2. The R package `alr3` contains a data set called `banknote`, consisting of (physical) measurements on 200 Swiss bank notes, 100 of which are genuine, while the other half is counterfeit. Load this data set (you might have to install the package) using the code below. Also run the cryptic third line - it will make your life a lot easier for the rest of the homework. This turns variable Y explicitly into a factor variable, i.e. makes it discrete. We will discuss this in the course material later in more detail.

```{r, message = FALSE}
# install.packages("alr3")
library(alr3) # if this throws an error of the form 'there is no package called alr3', uncomment the line above, run it once, then comment the line out again and run the code chunk again.
library("ggplot2", lib.loc="~/R/win-library/3.4")
data(banknote)
banknote$Y <- factor(banknote$Y)
```
3. Use one of our object inspecting functions and interpret the result in the data that you see.

```{r}
summary(banknote)
```

4. Use the package `ggplot2` to draw a barchart of Y (0 is genuine, 1 is counterfeit). Map Y to the fill color of the barchart. 

```{r }
caption <- paste(strwrap("Fig1: Bar chart showing number of different classes of banknotes in the data. From the barchart, we can see that we have a balanced dataset with 100 examples from both classes.", width = 95), collapse = "\n")
ggplot(banknote, aes(x=Y,color=Y)) + geom_bar(size=1)+
  labs(x = "Bottom", caption = caption) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right",
  plot.margin = margin(15, 15, 15, 15),
  plot.caption = element_text(hjust = 0))
ggplot(data = banknote, aes(x = Y, fill = Y)) + geom_bar() 
ggplot(banknote, aes(x = Diagonal)) + geom_histogram(binwidth = 0.3, aes(fill = Y))
```

5. Use the package `ggplot2` to draw a histogram of one of the variables in the dataset that shows a distinction between genuine and counterfeit banknotes. Use fill color to show this difference. Choose the binwidth such that there are no gaps in the middle range of the histogram.

```{r }
caption <- paste(strwrap("Fig2: Histogram of Length. The plot shows the histogram of Length coloured by class 0 or 1 where 0 means counterfeit and 1 means genuine note. From the plot we can see that in the middle region both classes have near 50% of notes in a given range of length. A few genuine notes have greatest length.", width = 95), collapse = "\n")

ggplot(banknote, aes(x=Length, color=Y)) + geom_histogram(fill="white",bins=20)+ geom_histogram(fill="white",alpha=0.5, position="identity",bins=20)+
  labs(x = "Length", caption = caption) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right",
  plot.margin = margin(15, 15, 15, 15),
  plot.caption = element_text(hjust = 0))
```

6. Use the package `ggplot2` to draw a scatterplot of two (continuous) measurements, color by Y. Try to find a pair of measurements that allow you to separate perfectly between genuine and counterfeit banknotes. 

```{r}
caption <- paste(strwrap("Fig3: Scatterplot of Length vs Diagonal. The plot shows how the two classes of banknotes varies in Length and Diagonal.Each point represents a banknote colored by its class. It appears that the two classes are linearly separable in these two dimentions.", width = 170), collapse = "\n")

ggplot(banknote, aes(x=Length, y=Diagonal, color=Y)) + geom_point(size=2)+
  labs(x = "Length", y = "Diagonal", caption = caption) +
  theme_bw(base_size = 11) +
  theme(legend.position = "right",
  plot.margin = margin(15, 15, 15, 15),
  plot.caption = element_text(hjust = 0))
```

7. For each of the three figures above, write a two-three sentence summary, describing the
    a. **structure of the plot**: what type of plot is it? Which variables are mapped to x, to y, and to the (fill) color?
    b. **main message of the plot**: what is your main finding, i.e. what do you want viewers to learn from the plot?
    c. **additional message**: point out anomalies or outliers, if there are any. 


Description of each plot is written below the plots.
